Estimating duration in the presence of misclassification

### 2024-02-08 ATACCC.tex
previous viral load ODE models
impact of factors on duration

Two major approaches have been taken to model SARS-CoV-2 viral load trajectories, both based on well-well-established methods within biostatisics.
The first approach is an ODE-based approach uses parameters representing biological properties of the virus and host~\autocite[e.g.:][]{ejimaEstimation,keVivo,kimQuantitative,goncalvesTiming,perelsonMechanistic}.
The model here is an application of a standard model for viral dynamics, used for many diseases\todo{cite history of viral load ODE models}.
Such models are rarely analytically-tractable and the system's evolution is computed using numerical methods.
% Both of these approaches have been used with other viruses and biomarkers.
The second approach, and the focus of this chapter, is using a piecewise linear model for the biomaker trajectory and random effects to allow for within-individual correlation.

@@

Different estimates were produced for vaccinated and unvaccinated individuals in the study.
However, there are many differences between individuals in these two groups.
For example, the vaccinated individuals are likely to be older and/or have co-morbidities because the roll-out of the vaccine in the UK prioritised these groups~\autocite{naoCovidVaccination}.
They were also infected by different variants because the vaccinated individuals were infected later.
All these factors could change the viral dynamics~\autocite{russellWithinhost} and hence the duration.
\todo{Do the impact of these factors agree with other estiamtes? Does immunity matter more than other factors?}
Therefore, the relationships shown here cannot be considered causal, or even generalisable to contexts where these other confounders are not present.

>>>>

Two major approaches have been taken to model SARS-CoV-2 viral load trajectories, both based on well-well-established methods within biostatisics.
The first approach is an ODE-based approach uses parameters representing biological properties of the virus and host~\autocite[e.g.:][]{ejimaEstimation,keVivo,kimQuantitative,goncalvesTiming,perelsonMechanistic}.
The model here is an application of a standard model for viral dynamics, used for many diseases~\autocite{perelsonMechanistic,hernandez-vargasInhost}.
Such models are rarely analytically-tractable and the system's evolution is computed using numerical methods.
% Both of these approaches have been used with other viruses and biomarkers.
The second approach, and the focus of this chapter, is using a piecewise linear model for the biomaker trajectory and random effects to allow for within-individual correlation.

@@

Different estimates were produced for vaccinated and unvaccinated individuals in the study.
However, there are many differences between individuals in these two groups.
For example, the vaccinated individuals are likely to be older and/or have co-morbidities because the roll-out of the vaccine in the UK prioritised these groups~\autocite{naoCovidVaccination}.
They were also infected by different variants because the vaccinated individuals were infected later.
All these factors could change the viral dynamics~\autocite{russellWithinhost,kisslerDynamicsVaccines} and hence the duration.
Therefore, the relationships shown here cannot be considered causal, or even generalisable to contexts where these other confounders are not present.
Despite these confounders, the result is in agreement with previous work showing that the Alpha and Delta variants of concern (compared to pre-Alpha variants) and being vaccinated are all associated with shorter durations of RT-PCR positivity, although the vaccination affect is larger~\autocite{kisslerDynamicsVaccines}.
